\subsection*{Purpose of this Document}
This document defines how ScienceEcosystem is financed, how financial independence is protected, and how long-term operational continuity is ensured. It complements the Governance Charter and Business Plan and is binding on future organisational evolution.

\section*{1. Funding Philosophy}

ScienceEcosystem is funded as public-interest infrastructure. Its financial model is designed to support maintenance, stewardship, and gradual improvement rather than growth, profit, or market dominance.

Financial sustainability is defined as the ability to operate reliably across decades without compromising independence, openness, or trust.

\section*{2. Prohibited Revenue Sources}

The following funding mechanisms are explicitly prohibited:
\begin{itemize}
    \item Paywalls or access fees for users
    \item Sale or monetisation of user data or behavioural analytics
    \item Advertising or sponsored ranking
    \item Venture capital or equity-based investment
    \item Exclusive commercial licensing of core infrastructure
\end{itemize}

These prohibitions are structural and may not be overturned without formal governance reform.

\section*{3. Permitted Funding Sources}

ScienceEcosystem may be funded through the following channels:

\subsection*{3.1 Public and Competitive Grants}
\begin{itemize}
    \item National research infrastructure grants
    \item International funding programmes (e.g. EU, UNESCO-aligned initiatives)
    \item Public digital infrastructure funding
\end{itemize}

Grants must not impose exclusive control, data enclosure, or directional influence.

\subsection*{3.2 Philanthropic and Foundation Funding}
\begin{itemize}
    \item Science, education, and open-knowledge foundations
    \item Long-term mission-aligned philanthropic donors
\end{itemize}

All donations above a defined threshold are publicly disclosed.

\subsection*{3.3 Voluntary Contributions}
\begin{itemize}
    \item Individual donations
    \item Institutional goodwill contributions without contractual control
\end{itemize}

Voluntary contributions do not confer governance influence.

\section*{4. Financial Transparency and Accountability}

ScienceEcosystem commits to:
\begin{itemize}
    \item Public annual financial summaries
    \item Independent audits once legally incorporated
    \item Clear separation of operational and reserve funds
\end{itemize}

All financial records are subject to governance oversight.

\section*{5. Cost Structure}

Core cost categories include:
\begin{itemize}
    \item Hosting and infrastructure
    \item Core technical maintenance
    \item Metadata integration and curation
    \item Governance and legal compliance
    \item Accessibility and documentation
\end{itemize}

Staffing remains minimal and role-based rather than growth-driven.

\section*{6. Endowment and Long-Term Stability}

The long-term objective is to establish an endowment that:
\begin{itemize}
    \item Covers baseline operational costs
    \item Reduces dependency on short-term grants
    \item Supports multi-decade continuity
\end{itemize}

Endowment governance follows the same independence constraints as operational funding.

\section*{7. Financial Risk Management}

Key risks include grant dependency and funding volatility. These are mitigated through:
\begin{itemize}
    \item Funding diversification
    \item Conservative cost control
    \item Multi-year planning horizons
\end{itemize}

\section*{8. Commitment}

ScienceEcosystem is financed to exist reliably, not to scale aggressively. Financial decisions must preserve legitimacy and trust across generations.

